Risks of Running Vulnerable Container Images for Data Analysis on HPC Clusters

### 2019-08-03 paper.tex
saturday

%%%% Commands %%%%
\newcommand{\todo}[1]{\color{red}\textbf{TODO:}#1\color{black}}
\newcommand{\note}[2]{\color{blue}Note: #1\color{black}}
\newcommand{\reprozip}[0]{ReproZip}
\newcommand{\TG}[1]{\color{blue}From Tristan: #1 \color{black}}

@@

However, two features of the Linux kernel, \textit{control groups} and \textit{namespaces},
are used by containers to mitigate these threats to a certain extent. Security
of containers is still most concerning issue. According to a Forrester survey~\cite{} done in January
2015, security is the
biggest concern while deploying containers.
Except web applications, containers are also popularly used for scientific computing
and data science in neuroscience. In fact, Singularity was developed for scientific applications
by keeping focus on HPC environments. This popularity containers lead us to the
primary research question of this work: \textit{what is the current security state of
container images that are deployed on HPC clusters}.

To answer this question, we take a real use case of containers into consideration.
The goal of this paper is to report all vulnerabilities that are present in
container images, specifically used by neuroimaging field, and access their
related risks for data analysis. Particularly, we aim to answer three primary
research questions:

\textit{RQ1} Are vulnerabilities present inside container images that are
deployed on HPC clusters?
\textit{RQ2} Can we get rid of these vulnerabilities by updating the
images?
\textit{RQ3} Can we get rid of these vulnerabilities by shrinking the
size of the image?

To answer those questions, we take an example of neuroscience field where
containers are used on HPC clusters.

The existing work on containers focuses mainly on the security of Docker
containers.
This focus is justified by the fact that, containers expose the host's resources
(e.g., file system/ IPC) to the guest system. This feature raises a confidentiality
threat for the applications running on the same host. Previous studies evaluated
security of Docker engine ~\cite{martin2018docker, sultan2019container, combe2016docker, bui2015analysis},
and have scanned vulnerabilities on Docker hub~\cite{Shu2017, gummaraju2015over}.
Also Singularity came up with Stools, which can be used for scanning Singularity images
for security and quality checks.
Here, we focus on the specific context of scientific data analysis on HPC clusters, taking
neuroimaging as an example.

\TG{Some notions that should be introduced in this section (incomplete):
vulnerability vs malware, paper suggested by Emad Shihab.}

@@

We used the 49 Boutiques applications published on Zenodo at the time of
the writing. These applications use 23 different container images,
including 7 Singularity images and 16 Docker images. \TG{You should add a
table showing the images available in each framework, with their base
image, DOIs, size, and number of files.}

@@

BIDS apps~\cite{gorgolewski2017bids} is a framework designed to share and execute neuroimaging
analysis pipelines. It improves usability, accesibility, and reproducibility
of different applications through containerization.
\TG{How many BIDS apps are there? Did you use them all?} Currently, there are 27 different container images that are used by
BIDS App. All these are Docker images.

@@

Out of the various container image scanning tools, we selected Anchore,
Vuls, Stools and ClairScanner for our experiments. \TG{Explain why you used
more than 1 tool, and why you used these ones speifically.}

@@

@@

Vuls is an open-source vulnerability scanner for Linux and FreeBSD, written in Go
language \TG{Why is it relevant that it's written in Go?}.
 It is an agentless vulnerability scanner and uses various
vulnerability databases \TG{You should list them and explain how they differ to 
the ones used by Anchore}. Using dynamic scanning technique, it offers
remote and local scan. Remote scan involves setting up only one machine, to
which other target servers are connected via SSH. Local scan involves running
the scan on the target server itself. Further, Vuls offer fast scan, fast root scan,
and deep scan.
Fast scan is performed without root privileges, whereas fast root scan is
performed with root privileges. The latter also uses checkrestart utility to detect processes that are updated but not yet restarted because if they are not
restarted, updates are not effective. So fast root scan also detects these pro-
cesses. Both fast scan and fast root scan are performed offline without Internet
access, and put almost no load on the scan target server. Deep scan uses
root privileges for scanning, and is used to get a more detailed scan. It checks
changelogs, which puts more load on the target server. As shown in Figure 4.3 \TG{There is no figure 4.3 or graph 4.5 here. Anyway it's not relevant 
to report results in this section. This paragraph looks like a copy paste from your previous report, it needs to be edited.}
it is clear that vulnerability detection time of Vuls is much better than An-
chore. Vuls tool uses dynamic scanning, so the scanning time does not depend
on the image size.
Graph 4.5 shows percentage of vulnerabilities that are
covered by Vuls and Anchore tools. This graph can be explained with the help
of List 4.4, which shows vulnerability databases that are referred, by both Vuls
and Anchore. Anchore covers less percentage of vulnerabilities as compared
to Vuls because Anchore relies mainly on vulnerability databases published by
Linux distributions, whereas Vuls additionally refers to another sources such as
databases from Japan, US-CERT, etc.

@@

Stools combines CoreOS's ClairScanner with the Continuous Integration (Travis and
Circle) for scanning Singularity images for vulnerabilities.
Initially, ClairScanner was intented to scan Docker Containers. This tool does not fit well
to scan Singularity images because of the different image formats. Docker layers
consist of layers, whereas Singluarity images consist of a single binary file.
However, TravisCI and CircleCI are commonly used for testing code, and these
services also offer running containers. So Stools combined both the concepts.
It involves spinning a ClairScanner container during testing and building a
Singularity image, and scanning it for filesystem before the image is finalized.
\TG{Add link and/or reference.}

>>>>

%%%% Commands %%%%
\newcommand{\todo}[1]{\color{red}\textbf{TODO:}#1\color{black}}
\newcommand{\note}[2]{\color{blue}Note: #1\color{black}}
\newcommand{\reprozip}[0]{ReproZip}
\newcommand{\TG}[1]{\color{blue}From Tristan: #1\color{black}}

@@

However, two features of the Linux kernel, \textit{control groups} and \textit{namespaces},
are used by containers to mitigate these threats to a certain extent. Security
of containers is still most concerning issue. According to a Forrester survey~\cite{} done in January
2015, security is the
biggest concern while deploying containers.
Except web applications, containers are also popularly used for scientific computing
and data science in neuroscience. In fact, Singularity was developed for scientific applications
by keeping focus on HPC environments. This popularity of containers lead us to the
primary research question of this work: \textit{what is the current security state of
container images that are deployed on HPC clusters}.

To answer that question, we scan all images that are used on Compute
Canada clusters by neuroscience field. Scanning refers to the practice of
collecting all information about the container image, investigating it
for vulnerabilities, and finally producing a
report to summarize scanning results. Scanning is broadly categorized into
two types: static scanning and dynamic scanning. The former scanning technique
involves extracting information of an image without running it into a container,
whereas the latter approach involves launching the container from the target image, and
then scanning the running container.
Here, we focus mainly on vulnerability scanning and not on malware.
Malware is a computer software that is specially designed to attack the target computer,
whereas vulnerability is a weakness in the developer's software that can be used
by the attacker to perform unathorized actions on the target computer.
In this paper, we report all vulnerabilities that are present in
container images, specifically used by neuroimaging field, and access their
related risks for data analysis. Particularly, we aim to answer three primary
research questions:

\textit{RQ1:} Are vulnerabilities present inside container images that are
deployed on HPC clusters?

\textit{RQ2:} Can we get rid of these vulnerabilities by updating the
images?

\textit{RQ3:} Can we get rid of these vulnerabilities by shrinking the
size of the image?

To answer those questions, we take an example of neuroscience field where
containers are used on Compute Canada HPC cluster.
We are not the first to work on vulnerabilities inside container images but
the existing research work on containers focuses mainly on the security of Docker
containers.
This focus is justified by the fact that, containers expose the host's resources
(e.g., file system/ IPC) to the guest system. This feature raises a confidentiality
threat for the applications running on the same host. Previous studies evaluated
security of Docker engine ~\cite{martin2018docker, sultan2019container, combe2016docker, bui2015analysis},
and have scanned vulnerabilities on Docker hub~\cite{Shu2017, gummaraju2015over}.
Also Singularity came up with Stools, which can be used for scanning Singularity images
for security and quality checks.
Here, we focus on the specific context of scientific data analysis on HPC clusters, taking
neuroimaging as an example.

\TG{paper suggested by Emad Shihab.}

@@

We used the 49 Boutiques applications published on Zenodo at the time of
the writing. These applications use 23 different container images,
including 7 Singularity images and 16 Docker images.

@@

BIDS apps~\cite{gorgolewski2017bids} is a framework designed to share and execute neuroimaging
analysis pipelines. It improves usability, accesibility, and reproducibility
of different applications through containerization.
Currently, there are 27 BIDS apps available and all are using different Docker images.

\begin{figure}
  \centering
  \includegraphics[width=.7\columnwidth]{Figures/bids_table.png}
  \caption{All Container images used by BIDS apps.}
  \label{fig:bids_apps}
\end{figure}

@@

Out of the various container image scanning tools available to scan Docker images,
we selected Anchore
and Vuls because these are the most mature scanning tools. To scan Singularity images
we use Stools, which is specially designed for Singularity images. This tool internally uses
ClairScanner tool so we included ClairScanner also in our experiements to compare
performance of both the tools.

@@

\begin{figure}
  \centering
  \includegraphics[width=.7\columnwidth]{Figures/anchore.png}
  \caption{Databases used by Anchore.}
        \label{fig:anchore_db}
\end{figure}

@@

Vuls is an open-source vulnerability scanner for Linux and FreeBSD.
It is an agentless vulnerability scanner and uses various
vulnerability databases \TG{You should list them and explain how they differ to 
the ones used by Anchore}. Using dynamic scanning technique, it offers
remote and local scan. Remote scan involves setting up only one machine, to
which other target servers are connected via SSH. Local scan involves running
the scan on the target server itself. Further, Vuls offer fast scan, fast root scan,
and deep scan.
Fast scan is performed without root privileges, whereas fast root scan is
performed with root privileges. The latter also uses checkrestart utility to 
detect processes that are updated but not yet restarted because if they are not
restarted, updates are not effective. So fast root scan also detects these pro-
cesses. Both fast scan and fast root scan are performed offline without Internet
access, and put almost no load on the scan target server. Deep scan uses
root privileges for scanning, and is used to get a more detailed scan. It checks
changelogs, which puts more load on the target server.

\begin{figure}
  \centering
  \includegraphics[width=.7\columnwidth]{Figures/vuls.png}
  \caption{Databases used by Vuls.}
	\label{fig:vuls_db}
\end{figure}

@@

\href{https://github.com/singularityhub/stools}{Stools} combines CoreOS's ClairScanner with the Continuous Integration (Travis and
Circle) for scanning Singularity images for vulnerabilities.
Initially, ClairScanner was intented to scan Docker Containers. This tool does not fit well
to scan Singularity images because of the different image formats. Docker layers
consist of layers, whereas Singluarity images consist of a single binary file.
However, TravisCI and CircleCI are commonly used for testing code, and these
services also offer running containers. So Stools combined both the concepts.
It involves spinning a ClairScanner container during testing and building a
Singularity image, and scanning it for filesystem before the image is finalized.
